---
title: "ST516: Computational Sttistics Exam (Part B)"
author: "Katrine Eriksen and Katrine Bach"
date: "2 juni 2016"
output: pdf_document
---

#Task 4
###Standard Normal Distribution:
###Matrix regression:
###Linear regression: 
###Residual standard error:
###$R^2$:
###$R_{adj}^2$:
###F-Statistics:

***

#Task 5
**Theory**
##Standard Normal Density
##Naive density estimation
The Naive kernel method uses each point of estimation $x$ as the center og the bin of width $2h$. 
### Kernel density estimation
The Kernel density estimation is an esppraoch that is rooted in the  histogram methodology. The basic idea is to estimate the density function at a point $x$ using neighboring observations. To express it more transparently, we consider the weight function 
$$55$$
which is called the kernel weight. The kernel estimate of $f(x)$ is defined as
$$55$$
###Naive estimator:
###Gaussian Kernel 
###Silverman's suggestion for Gaussian method: 
###Density:

Here we apply our function densi that estimates the density using the naive estimator ot thre gaussian kernel. First we run it foe the Kernel density estimation without specifying $d$ or $h$.
```{r}
densi <- function(x,d=NULL,h=NULL, method="naive"){
n<-length(x)

if(length(h)==0){
  if (method=="kernel"){
    h <- 0.9*min(c(IQR(x)/1.34,sd(x)))*n^(-1/5)
  }else{
    h<- (max(x)-min(x))/(1+log2(n))
  }
  #Make warning with other methods
}else{
  h <- h
}

if(length(d)==0){
  minimum<-densi(x,min(x),h,method)
  maksimum<-densi(x,max(x),h, method)

  quantile1<-quantile(x,names=FALSE)[2]
  q1<-densi(x,quantile1,h,method)

  quantile3<-quantile(x,names=FALSE)[4]
  q3<-densi(x,quantile3,h,method)

  medians <- densi(x,median(x),h,method)
  means <- densi(x,mean(x),h,method)

  smoke <- matrix(c(min(x),minimum,quantile1,q1,median(x),medians,mean(x),means,quantile3,q3,max(x),maksimum), ncol = 2, byrow = TRUE)
  colnames(smoke) <- c("x","y")
  rownames(smoke) <- c("Min","1st Quantile","Median", "Mean", "3dr Quantile", "Max")
  smoke <- as.table(smoke)
  print(sprintf("bandwidth: %f", h))
  return(smoke)

}else{
  d=d
}

if(method=="kernel"){
  kern<-0
  for (i in 1:n) {
    kern[i] <-(1/h)*dnorm((d-x[i])/h)
  }
  summ<- sum(kern)
  f<-(1/n)*summ
  return(f)
}else{
  nav <- 0
  for(i in 1:n){
    if(abs((d-x[i])/h)<1){
      w <- 1/2
  }else{
      w <- 0
  }
    nav[i] <-(1/h)*w
  }
  summ <- sum(nav)
  f <- (1/n)*summ
  return(f)
}

}
densi(faithful$eruptions, method="kernel")



```

**Forklar hvad værdierne viser**

Now we run it again using the Naive density estimation and again without specifying $d$ and $h$. 

```{r}
densi(faithful$eruptions)

```

**Forklar noget om værdierne** 

Now we use our function densiplot, that calls our density function and creates a density plot based on the specified method. We start by showing the plot for the Kernel denisty estimation. 
```{r}
densiplot <- function(x,n=500, method="naive", from= min(x)-(sd(x)/3), to=max(x)+(sd(x)/3)){
  happy<-c()
  for (d in seq(from,to,length.out = n)) {
    happy1<-densi(x,d=d,method=method)
    happy<-c(happy,happy1)
  }
  plot(happy)
}
densiplot(faithful$eruptions, method="kernel")

```

Now we run it again using the Naive density estimation and again without specifying $d$ and $h$. 

```{r}
densiplot(faithful$eruptions)

```


And now we want to se how our densiplot function looks againts the already build-in function *density()*

```{r}
plot(density(faithful$eruptions))
density(faithful$eruptions)

```

Her kan man også sammenligne tabellen med hinanden. Husk at lav funktionsbeskrivelser til densi() og densiplot()!!!!

***

#Task 6
###Markov chain:
###Transition Probability:

***

#Task 7
###Beta Distribution:
###Metropolis-Hastings sampler: 
###Rejection rate: 



***
